---
title: "PA1"
author: "Adam McCullough"
date: "August 23, 2016"
output: html_document
---
##Step 1
### Load and Process Data

Load the data(note: the data file was downloaded and extracted to the working directory)
```{r , echo=TRUE}
activity<-read.csv("activity.csv")

```

##Step 2
###Initial interpretation

Sum the number of steps by day and create a histogram of the output
```{r , echo=TRUE}
library(dplyr)
sum1<- na.omit(activity) %>% group_by(date) %>% summarize(Steps=sum(steps))
hist(sum1$Steps, breaks = 10, xlab = "Steps", main = "Histogram of Steps by Day")
meanbyday<-mean(sum1$Steps)
meanbyday
medianbydate<-median(sum1$Steps)
medianbydate
```

##Step 3
###Average daily activity

Need to group and average steps by time interval, then plot a line graph
```{r, echo=TRUE}
library(dplyr)
sum2<- na.omit(activity) %>% group_by(interval) %>% summarize(Steps=mean(steps))
plot(sum2$interval,sum2$Steps, type = "l",xlab="Interval",ylab = "Steps",main = "Steps by Time Interval")

```

Now determine and output the interval with the max average value across all days
```{r setup, echo=TRUE}
subset(sum2, sum2$Steps == max(sum2$Steps))

```

##Step 4 
###Imputing missing values

Calculate the total number of rows with missing values
```{r, echo=TRUE}
sum(is.na(activity))
```

Fill in these missing values with the mean value for that time interval across all other days
```{r, echo=TRUE}
activity2<-activity
activity2$steps[is.na(activity2$steps)] <- sum2$Steps[match(activity2$interval[is.na(activity2$steps)],sum2$interval)]
```

First summarize by day then, create histogram and output mean and median of imputted data
```{r, echo=TRUE}
sum3<- activity2 %>% group_by(date) %>% summarize(steps=sum(steps))
hist(sum3$steps, breaks = 10, xlab = "Steps", main = "Histogram of Steps by Day")
meanbyday2<-mean(sum3$steps)
meanbyday2
medianbydate2<-median(sum3$steps)
medianbydate2
```
##Step 5
###Weekdays vs Weekends

Determine whether each date is a weekday or weekend
```{r, echo=TRUE}
activity2$weekday<-weekdays(as.POSIXlt(activity2$date))
 activity2$wd<-ifelse(activity2$weekday %in% c("Sunday","Saturday"), "Weekend", "Weekday")
```

Plot two graphs representing the average steps by time interval grouped seperately by Weekday and Weekend
```{r, echo=TRUE}
par(mfrow=c(2,1))
weekday<-activity2[activity2$wd=="Weekday",]
plot1<- weekday %>% group_by(interval) %>% summarize(Steps=mean(steps))
weekend<-activity2[activity2$wd=="Weekend",]
plot2<- weekend %>% group_by(interval) %>% summarize(Steps=mean(steps))
plot(plot1$interval,plot1$Steps, type = "l",xlab="Interval",ylab = "Steps",main = "Steps by Time Interval- Weekday")
plot(plot2$interval,plot2$Steps, type = "l",xlab="Interval",ylab = "Steps",main = "Steps by Time Interval- Weekend")

```